More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-09-28 2016-Weak-FH-Bisim/WeakBisim.tex
partial

To have some practical interest, it is important to know when this
algorithm terminates. The following theorem shows that an open-pNet
with finite synchronisation sets, finitely many leaves and
holes, and each pLTS at leaves having a finite number of states and
(symbolic) transitions, has a finite automaton:

\begin{theorem}[Finiteness of open-automata.]\\
Given an open pNet $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ with leaves $pLTS_i^{i\in L}$ and holes $Hole_j^{j\in
  J}$, if the sets $L$ and $J$ are finite, if the synchronisation vectors of all pNets 
  included in  $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ 
  are finite, and if
$\forall i \in L.\, finite{(states(pLTS_i))} \text{ and } pLTS_i$
has a finite number of state variables, then Algorithm 1 terminates
and produces an open automaton 
$\mathcal{T}$ with finitely many states and transitions.

@@

% $$ finite{(L)} \land finite{(J)} \land \forall i \in L finite{(\{s_i\})}
%   \to finite{(\mathcal{T})}$$
\end{theorem}

\section{Tooling}

\ERIC{
=> Context VCE, ref to FASE paper\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule Tr2, the resulting OT is built from the premisses, without explicitely constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone'' operations, using structured variable names ensuring uniqueness of fresh names, but also reasonable readability of these names for debugging purposes.\\
=> Satisfiability check is done only at toplevel of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentialy reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
=> Simplification is not yet implemented. It is not strictly required for the Open Automaton construction, but it will be critical later for predicate comparison in the bisimulation algorithm.}

@@

Weak symbolic bisimulation was introduced to relate transition systems
that have indistinguishable behaviour, with respect to some definition
of \emph{invisible actions} that are considered internal to some
subsystem, and consequently cannot be observed, nor used for
synchronisation with their context.
As the notions of invisible action varies in different contexts,
e.g. $tau$ in CCS, and $i$ in Lotos, we will rely here on some set of
\emph{invisible actions} depending on a specific action algebra.

\TODO{we should understand if it make sense to distinguish invisible
  from synchronised actions: in the Forte paper we wrote ``using  as 
\emph{invisible actions} a subset of the
\emph{synchronised actions} defined in Section~\ref{section:pnets}''}.

@@

\noindent
Let $A_1$ and $A_2$ be open automata, and $mathcal{R}$ a relation over
$S_1$ and $S_2$, as in Definition \ref{def-FH-bisim}.

@@

@@

       It is trivial to check reflexivity and symmetry. Here we focus on the
transitivity. 
To prove transitivity of strong FH-bisimulation on pNets it is sufficient to prove 
transitivity of the strong FH-bisimulation on states. Consider 3 open automata 
$\mathcal{T}_1$, $\mathcal{T}_2$, $\mathcal{T}_3$ and states $s^1$, $s^2$, $s^3$ in those 
automata\footnote{We omit the constraints stating that each $s^i_x$ is in the states of 
$\mathcal{T}_i$ for the sake of readability}.
Suppose we have $\mathcal{R}$ an FH-bisimulation relation between states of 
$\mathcal{T}_1$ and of  $\mathcal{T}_2$; members of $\mathcal{R}$ are of the form 
$(s^1,s^2|\Pred)$.
Suppose we also  have $\mathcal{R}'$ an FH-bisimulation relation between states of 
$\mathcal{T}_2$ and of  $\mathcal{T}_3$; members of $\mathcal{R}2$ are of the form 
$(s^2,s^3|\Pred\,')$.
We use $\fv$ a function returning the free variables of a term/automaton.

>>>>

Given an open-pNet
with finite synchronisation sets, finitely many leaves and
holes, and each pLTS at leaves having a finite number of states and
(symbolic) transitions, has a finite automaton (see \ref{Forte}).

  \TODO{Eric: what is ``elimination of intermediate variables ?'' it
    transforms the predicates, so I think it is not any kind of
    ``equivalence'' transformation, rather some sort of
    abstraction... Define... Does it change the satisfiability ?}

%% \begin{theorem}[Finiteness of open-automata.]\\
%% Given an open pNet $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ with leaves $pLTS_i^{i\in L}$ and holes $Hole_j^{j\in
%%   J}$, if the sets $L$ and $J$ are finite, if the synchronisation vectors of all pNets 
%%   included in  $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ 
%%   are finite, and if
%% $\forall i \in L.\, finite{(states(pLTS_i))} \text{ and } pLTS_i$
%% has a finite number of state variables, then Algorithm 1 terminates
%% and produces an open automaton 
%% $\mathcal{T}$ with finitely many states and transitions.

@@

% $$ finite{(L)} \land finite{(J)} \land \forall i \in L finite{(\{s_i\})}
%   \to finite{(\mathcal{T})}$$
% \end{theorem}

@@

Weak symbolic bisimulation was introduced to relate transition systems
that have indistinguishable behaviour, with respect to some definition
of \emph{internal actions} that are considered local to some
subsystem, and consequently cannot be observed, nor used for
synchronisation with their context.
As the notions of invisible action varies in different contexts,
e.g. $tau$ in CCS, and $i$ in Lotos, we will rely here on some set of
\emph{internal actions} depending on a specific action
algebra. Naturely, an internal action cannot be synchronised with
actions of other systems in its environment.

%% \TODO{we should understand if it make sense to distinguish invisible
%%   from synchronised actions: in the Forte paper we wrote ``using  as 
%% \emph{invisible actions} a subset of the
%% \emph{synchronised actions} defined in Section~\ref{section:pnets}''}.

@@

\noindent
Let $A_1$ and $A_2$ be open automata, and $\mathcal{R}$ a relation over
$S_1$ and $S_2$, as in Definition \ref{def-FH-bisim}.

@@

\ERIC{
=> Context VCE, ref to FASE paper\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule Tr2, the resulting OT is built from the premisses, without explicitely constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone'' operations, using structured variable names ensuring uniqueness of fresh names, but also reasonable readability of these names for debugging purposes.\\
=> Satisfiability check is done only at toplevel of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentialy reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
=> Simplification is mostly implemented. It is not strictly required for the Open Automaton construction, but it will be critical later for predicate comparison in the bisimulation algorithm.}

@@

       It is trivial to check reflexivity and symmetry. Here we focus on the
transitivity. 
To prove transitivity of strong FH-bisimulation on pNets it is sufficient to prove 
transitivity of the strong FH-bisimulation on states of their Open Automata. Consider 3 open automata 
$\mathcal{T}_1$, $\mathcal{T}_2$, $\mathcal{T}_3$ and states $s^1$, $s^2$, $s^3$ in those 
automata\footnote{We omit the constraints stating that each $s^i_x$ is in the states of 
$\mathcal{T}_i$ for the sake of readability}.
Suppose we have $\mathcal{R}$ an FH-bisimulation relation between states of 
$\mathcal{T}_1$ and of  $\mathcal{T}_2$; members of $\mathcal{R}$ are of the form 
$(s^1,s^2|\Pred)$.
Suppose we also  have $\mathcal{R}'$ an FH-bisimulation relation between states of 
$\mathcal{T}_2$ and of  $\mathcal{T}_3$; members of $\mathcal{R}2$ are of the form 
$(s^2,s^3|\Pred\,')$.
We use $\fv$ a function returning the free variables of a term/automaton.
